---
title: "Prediction Assignment Writeup"
author: "Luying Feng"
date: "05/09/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Summary

The goal of the project is to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants to predict the manner in which they did the exercise. This is the "classe" variable in the training set.
The report describes the model, cross validation, expected out of sample error, and why the choices are made. 
The prediction model is also used to predict 20 different test cases.

The data for this project come from this source: http://web.archive.org/web/20161224072740/http:/groupware.les.inf.puc-rio.br/har.

# Import data
```{r ImData}
getwd()
setwd("...")

fileUrlTrain<-"https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv "
download.file(fileUrlTrain,destfile = ".\\Training.csv",mode='wb',cacheOK = FALSE)

fileUrlTest<-"https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv "
download.file(fileUrlTest,destfile = ".\\Testing.csv",mode='wb',cacheOK = FALSE)

Training<-read.csv(".\\Training.csv")
Testing<-read.csv(".\\Testing.csv")
    
```

