---
title: "Prediction Assignment Writeup"
author: "Luying Feng"
date: "05/09/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Summary

The goal of the project is to explore the concept of human activity recognition, the project uses data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants to predict the manner in which they did the exercise. This is the "classe" variable in the training set.
The report describes the model, cross validation, expected out of sample error, and why the choices are made. 
The prediction model is also used to predict 20 different test cases.

The data for this project come from this source: http://web.archive.org/web/20161224072740/http:/groupware.les.inf.puc-rio.br/har.

Six young health participants were asked to perform one set of 10 repetitions of the Unilateral Dumbbell Biceps Curl in five different fashions and classed as below: 

    A - exactly according to the specification  
    B - throwing the elbows to the front 
    c - lifting the dumbbell only halfway
    D - lowering the dumbbell only halfway
    E - throwing the hips to the front



# Import and clean data
```{r ImData}

UrlTrain<-"https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv "

UrlTest<-"https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv "

Training<-read.csv(url(UrlTrain))
Testing<-read.csv(url(UrlTest))

# Remove all 0 columns
Training<-Training[,colSums(is.na(Training)) == 0]
Testing <-Testing[,colSums(is.na(Testing)) == 0]

# Remove irrelavent columns such as user_name, raw_timestamp_part_1, raw_timestamp_part_2, cvtd_timestamp, new_window,num_window, column[1:7]
Training   <-Training[,-c(1:7)]
Testing <-Testing[,-c(1:7)]

# Remove dependent columns,
drops<-"kurtosis|skewness|max|min|amplitude|avg|stddev|var"
Training<-Training[,-grep(drops,colnames(Training))]

table(Training$classe)
plot(as.factor(Training$classe),main="Classe in the whole training set",xlab="classe",ylab="frequency")
    
```

# Load libraries and creat partitions of the training data into training set and test test

```{r models}
library(caret)
library(randomForest)
library(rpart)
library(rpart.plot)

# Partitions into training set and test set, p=3/4

inTrain = createDataPartition(Training$classe, p = 3/4,list=FALSE)
train = Training[inTrain, ]
test = Training[-inTrain, ]
    
```
# Decision tree 

```{r Dtree}
mod_dt<-rpart(classe~.,data=train,method="class")
pred_dt<-predict(mod_dt,test,type="class")

rpart.plot(mod_dt,cex=0.2)

confusionMatrix(pred_dt, as.factor(test$classe))

```
# Random forest 
```{r forest}
mod_rf<-randomForest(as.factor(classe) ~.,data=train)

pred_rf<-predict(mod_rf,test,type="class")

confusionMatrix(pred_rf, as.factor(test$classe))

```
# Comparing results
Random forest perform better than decision tree. The accuracy of random forest is 0.995 ( 95% CI : (0.993, 0.997)) while decision tree only achieve accuracy of 0.7429 ( 95% CI : (0.7304, 0.755)). The random forest model will be used to predict the 20 different test cases

# Use random forest model to predict the 20 different cases
The prediction results of the 20 different test cases are shown as below.
```{r predict}
new_pred<-predict(mod_rf,Testing,type="class")
new_pred
plot(new_pred,main="Prediction of classe in the 20 test set",xlab="classe",ylab="frequency")
```

# write to file
rmarkdown::render("Assignment.Rmd")